---
title: "second Task"
author: "Alejandro Vergara Rincon 81190"
date: "2023-08-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

After making the pertinent connections with Github, there will be the DESCRIPTION file and we proceed to fill in the required spaces, in this case it is filled like this:

Package: test1

Type: Package

Title: Package for task 2 development.

Version: 0.1.0

Author: Jose Alejandro Vergara 81190 Maintainer:

The package maintainer [Josea.vergararar\@ecci.edu.co](mailto:Josea.vergararar@ecci.edu.co){.email}

Description: This code is used to solve the data transformation exercises:<https://r4ds.had.co.nz/transform.html#transform>

License: "use_mit_license"

Encoding: UTF-8

LazyData: true

Imports: dplyr, tidyr, knitr

RoxygenNote: 7.2.3

## 2

```{r}
SYA <- function(x) {
  library(nycflights13)
  library(tidyverse)
  library(knitr)
  if (x == 1) {
    return(Ex1())
  } else if (x == 2) {
    return(Ex2())
  } else if (x == 3) {
    return(Ex3())}
  else if (x == 4) {
  return(Ex4())}
  else if (x == 5) {
    return(Ex5())}
  else if (x == 6) {
    return(Ex6())}
  else {

    print("It is not possible to perform this operation")
  }
}
```

This code defines an R function named **`SYA`** (search_your_asnwer). This function takes a single argument **`x`**, which is used to determine which specific task or operation to perform. Let's break down the code step by step:

1.  **Loading Libraries**: The function starts by loading several R packages: **`nycflights13`**, **`tidyverse`**, and **`knitr`**. These packages provide various functions and tools for data manipulation, analysis, and visualization.

2.  **Conditional Execution**: The function uses a series of **`if`** and **`else if`** statements to determine what action to take based on the value of the input **`x`**.

    -   If **`x`** is equal to 1, the function will return the result of the **`Ex1()`** function.

    -   If **`x`** is equal to 2, the function will return the result of the **`Ex2()`** function.

    -   If **`x`** is equal to 3, the function will return the result of the **`Ex3()`** function.

    -   If **`x`** is equal to 4, the function will return the result of the **`Ex4()`** function.

    -   If **`x`** is equal to 5, the function will return the result of the **`Ex5()`** function.

    -   If **`x`** is equal to 6, the function will return the result of the **`Ex6()`** function.

    -   If none of the above conditions are met, a message "It is not possible to perform this operation" will be printed.

3.  **Function Execution**: The function will execute the specified sub-function (e.g., **`Ex1()`**, **`Ex2()`**, etc.) based on the value of **`x`** that was passed as an argument. Each **`Ex1`**, **`Ex2, etc.`** corresponds to applying the solutions developed in the previous task, therefore the solution code for each one is identical to the solution in the first assignment.

    ## Example

```{r echo=FALSE}
Ex1 <- function(x){

library(nycflights13)
library(tidyverse)
  library(knitr)
exercise <- nycflights13::flights

Data_1 <- filter(exercise, arr_delay >= "2")
Data_2 <- filter(exercise, dest == "IAH" | dest == "HOU")
cat("\n")
cat("5.2.4 Exercises:")
cat("\n")
cat("1\n")

cat("#The first step is to load the nycflights13 and tidyverse libraries:\n")

cat("\n")
cat("\n")

cat("library(nycflights13)\n")
cat("library(tidyverse)\n")
cat("library(knitr)\n")

cat("\n")
cat("\n")
cat(
"#Then the flight data is assigned to a called variable,
 prin this case the variable exercise This allows you to use the variable exercise to reference
 the data in the code:\n")

cat("\n")
cat("\n")

cat("exercise <- nycflights13::flights\n")

cat("\n")
cat("\n")

cat("Data_1 <- filter(exercise, arr_delay >= 2)\n")
cat("\n")
cat("\n")
cat(
"#A new dataset named Data_1 is created by using the filter() function,
which is part of the dplyr package within tidyverse. This function allows you to filter
the data according to a given condition.\n")
cat("\n")
cat("\n")

print(kable(Data_1[1:10, c(1, 14, 13, 9)],
            caption = "In this table, you can see information about the flights.",
            align = "c"))
cat("\n")
cat("2\n")


cat("Data_2 <- filter(exercise, dest == \"IAH\" | dest == \"HOU\")\n")

cat("\n")

cat(
"# In this code, we use the OR operator (\\|) to filter out flights that are bound for Houston,
either IAH or HOU. The condition dest == \"IAH\" \\| dest == \"HOU\" ensures that flights that meet
either of these two options are included.\n\n")
cat("\n")

print(kable(Data_2[1:15, c(1, 14, 13, 9)],
            caption = "In this table, you can see information about the flights and destinations.",
            align = "c"))}
```

Here it is evident how by putting SYA(1) it extracts what was done in exercise 1 with a short explanation

```{r}
library("test1")
SYA(1)
```
